Computing Entropies With Nested Sampling

### 2017-06-19 paper/ms.tex
A bit more detail in the algorithm section

We would like to calculate the expected value of a log-probability:

See Algorithm~\ref{alg:algorithm} for a step-by-step description of the
algorithm. The idea is to generate a reference particle from the distribution
whose entropy is required, and then use Nested Sampling to measure the
log-probability of a small region near that particle.

The algorithm is written in terms of an unknown quantity $\x$
whose entropy is required. In applications, $\x$ may be parameters,
a subset of the parameters, data, parameters and data jointly, or
whatever.

@@

\While{more iterations desired}
    \State $k \leftarrow 0$
            \Comment{Initialise counter}
    \State Generate $\x_{\rm ref}$ from $p(\x)$
            \Comment{Generate a reference point}
\color{green}
    \State Generate $\left\{\x_i\right\}_{i=1}^N$ from $p(\x)$
            \Comment{Generate initial NS particles}
    \State Calculate $d_i \leftarrow d(\x_i; \x_{\rm ref})$ for all $i$
            \Comment{Calculate distance of each particle from the reference
                     point}
    \State $i^* \leftarrow \textnormal{argmin}(\lambda i \rightarrow d_i)$
            \Comment{Find the worst particle (greatest distance)}
    \State $d_{\rm max} \leftarrow d_{i^*}$
            \Comment{Find the greatest distance}
    \While{$d_{\rm max} > L$}
        \State Replace $\x_{i^*}$ with $\x_{\rm new}$ from
               $p\left(\x | (d(\x) < d_{\rm max})\right)$
            \Comment{Replace worst particle}
        \State Calculate $d_{i^*} \leftarrow d(\x_{\rm new}; \x_{\rm ref})$
            \Comment{Calculate distance of new particle from reference point}
        \State $i^* \leftarrow \textnormal{argmin}(\lambda i \rightarrow d_i)$
                \Comment{Find the worst particle (greatest distance)}
        \State $d_{\rm max} \leftarrow d_{i^*}$
                \Comment{Find the greatest distance}
        \State $k \leftarrow k+1$
            \Comment{Increment counter $k$}
    \EndWhile
\color{black}
    \State $\widehat{\mathbf{h}} \leftarrow \widehat{\mathbf{h}} + [k/N]$
            \Comment{Append latest estimate to results}
\EndWhile
\State $\widehat{h_{\rm final}} \leftarrow \frac{1}{\rm num\_iterations} \sum \widehat{\mathbf{h}}$
            \Comment{Average results}
\end{algorithmic}
\caption{Calculates $H(\x) = -\int p(\x) \log p(\x) \, d\x$
         without evaluating $\log p(\x)$. Technically, what it actually
         estimates is $\int f(\x) \int f(\x')
            \log \left[ P(d(\x', \x) < L | \x) \right]
                        \, d\x' \, d\x$, the expected value of the
        log-probability of a small region near $\x$.
        The part highlighted in green is standard Nested Sampling.
        \label{alg:algorithm}}
\end{algorithm}

>>>>

The above insight, that the number of NS discarded points with likelihood
below $\ell$ has expected value $-\ln X(\ell)$, is the basis of the algorithm.
If we wish to measure a log-probability $\ln X(\ell)$, we can use NS to do it.
See Algorithm~\ref{alg:algorithm} for a step-by-step description of the
algorithm. The algorithm is written in terms of an unknown quantity $\x$
whose entropy is required. In applications, $\x$ may be parameters,
a subset of the parameters, data, some function of the data,
parameters and data jointly, or whatever.

The idea is to generate a reference particle $\x_{\rm ref}$
from the distribution $p(\x)$
whose entropy is required. Then, a Nested Sampling run evolves a set of
particles $\{\x\}_{i=1}^N$, initially representing $p(\x)$, is evolved
towards $\x_{\rm ref}$ in order to measure the log-probability near
$\x_{\rm ref}$.

@@

\While{more iterations desired}
    \State $k \leftarrow 0$
            \Comment{Initialise counter}
    \State Generate $\x_{\rm ref}$ from $p(\x)$
            \Comment{Generate a reference point}
\color{green}
    \State Generate $\left\{\x_i\right\}_{i=1}^N$ from $p(\x)$
            \Comment{Generate initial NS particles}
    \State Calculate $d_i \leftarrow d(\x_i; \x_{\rm ref})$ for all $i$
            \Comment{Calculate distance of each particle from the reference
                     point}
    \State $i^* \leftarrow \textnormal{argmin}(\lambda i \rightarrow d_i)$
            \Comment{Find the worst particle (greatest distance)}
    \State $d_{\rm max} \leftarrow d_{i^*}$
            \Comment{Find the greatest distance}
    \While{$d_{\rm max} > L$}
        \State Replace $\x_{i^*}$ with $\x_{\rm new}$ from
               $p\left(\x | (d(\x) < d_{\rm max})\right)$
            \Comment{Replace worst particle}
        \State Calculate $d_{i^*} \leftarrow d(\x_{\rm new}; \x_{\rm ref})$
            \Comment{Calculate distance of new particle from reference point}
        \State $i^* \leftarrow \textnormal{argmin}(\lambda i \rightarrow d_i)$
                \Comment{Find the worst particle (greatest distance)}
        \State $d_{\rm max} \leftarrow d_{i^*}$
                \Comment{Find the greatest distance}
        \State $k \leftarrow k+1$
            \Comment{Increment counter $k$}
    \EndWhile
\color{black}
    \State $\widehat{\mathbf{h}} \leftarrow \widehat{\mathbf{h}} + [k/N]$
            \Comment{Append latest estimate to results}
\EndWhile
\State $\widehat{h_{\rm final}} \leftarrow \frac{1}{\rm num\_iterations} \sum \widehat{\mathbf{h}}$
            \Comment{Average results}
\end{algorithmic}
\caption{Calculates $H(\x) = -\int p(\x) \log p(\x) \, d\x$
         without evaluating $\log p(\x)$. Technically, what it actually
         estimates is $\int f(\x) \int f(\x')
            \log \left[ P(d(\x', \x) < L | \x) \right]
                        \, d\x' \, d\x$, the expected value of the
        log-probability of a small region near $\x$.
        The part highlighted in green is standard Nested Sampling
        with quasi-prior $p(\x)$ and quasi-likelihood given by minus a
        distance function $d(\x', \x)$.
        \label{alg:algorithm}}
\end{algorithm}
